---
title: "ML Project"
author: "Mark"
date: "December 25, 2015"
output: html_document
---

This report describes the analysis done to predict correct exercise technique based on the data set: http://groupware.les.inf.puc-rio.br/har.

First load libraries:
```{r, echo=FALSE}
library(caret)
library(randomForest)
```

Load the training and testing data:
```{r, cache=TRUE}
training = read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"),header=TRUE,stringsAsFactors=FALSE)
testing = read.csv(url("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"),header=TRUE,stringsAsFactors=FALSE)
```

For cross validation I reserved 25% of the training samples to evaluate the accuracy of the model.
Partition the training data set in two sets, one for train and one for evaluation:
```{r, cache=TRUE}
inTrain <- createDataPartition(y=training$classe,p=0.75,list=FALSE)
ptraining1 <- training[inTrain,]
ptesting1 <- training[-inTrain,]

```
To make the run time of the predictor reasonable I prunned the data by
removing all columns from the partitions that are:  

* associated with user name or time. 
* not in the testing set.  
* have NA in every row.  
```{r, cache=TRUE}
ptraining <- subset(ptraining1, select=-c(user_name,X,raw_timestamp_part_1,raw_timestamp_part_2,cvtd_timestamp,new_window,num_window,
                                       kurtosis_roll_belt,
                                       kurtosis_picth_belt,
                                       kurtosis_yaw_belt,
                                       skewness_roll_belt,
                                       skewness_roll_belt.1,
                                       skewness_yaw_belt,
                                       max_yaw_belt,
                                       min_yaw_belt,
                                       amplitude_yaw_belt,
                                       kurtosis_roll_arm,
                                       kurtosis_picth_arm,
                                       kurtosis_yaw_arm,
                                       skewness_roll_arm,
                                       skewness_pitch_arm,
                                       skewness_yaw_arm,
                                       kurtosis_roll_dumbbell,
                                       kurtosis_picth_dumbbell,
                                       kurtosis_yaw_dumbbell,
                                       skewness_roll_dumbbell,
                                       skewness_pitch_dumbbell,
                                       skewness_yaw_dumbbell,
                                       max_yaw_dumbbell,
                                       min_yaw_dumbbell,
                                       amplitude_yaw_dumbbell,
                                       kurtosis_roll_forearm,
                                       kurtosis_picth_forearm,
                                       kurtosis_roll_forearm,
                                       kurtosis_yaw_forearm,
                                       skewness_roll_forearm,
                                       skewness_pitch_forearm,
                                       skewness_yaw_forearm,
                                       max_yaw_forearm,
                                       min_yaw_forearm,
                                       amplitude_yaw_forearm
                                       ))
ptesting <- subset(ptesting1, select=-c(user_name,X,raw_timestamp_part_1,raw_timestamp_part_2,cvtd_timestamp,new_window,num_window,
                                       kurtosis_roll_belt,
                                       kurtosis_picth_belt,
                                       kurtosis_yaw_belt,
                                       skewness_roll_belt,
                                       skewness_roll_belt.1,
                                       skewness_yaw_belt,
                                       max_yaw_belt,
                                       min_yaw_belt,
                                       amplitude_yaw_belt,
                                       kurtosis_roll_arm,
                                       kurtosis_picth_arm,
                                       kurtosis_yaw_arm,
                                       skewness_roll_arm,
                                       skewness_pitch_arm,
                                       skewness_yaw_arm,
                                       kurtosis_roll_dumbbell,
                                       kurtosis_picth_dumbbell,
                                       kurtosis_yaw_dumbbell,
                                       skewness_roll_dumbbell,
                                       skewness_pitch_dumbbell,
                                       skewness_yaw_dumbbell,
                                       max_yaw_dumbbell,
                                       min_yaw_dumbbell,
                                       amplitude_yaw_dumbbell,
                                       kurtosis_roll_forearm,
                                       kurtosis_picth_forearm,
                                       kurtosis_roll_forearm,
                                       kurtosis_yaw_forearm,
                                       skewness_roll_forearm,
                                       skewness_pitch_forearm,
                                       skewness_yaw_forearm,
                                       max_yaw_forearm,
                                       min_yaw_forearm,
                                       amplitude_yaw_forearm
))
ptraining <- ptraining[,colSums(is.na(ptraining))==0]
ptesting <- ptesting[,colSums(is.na(ptesting))==0]
```

A model was created with train function using the 'random forest' method to explore optimum settings. The optimium settings where with mtry=2.
Create a random tree model with these parameters:
```{r, cache=TRUE}
model <- randomForest(as.factor(classe) ~ ., data=ptraining, mtry=2,importance=TRUE)
```

For the training partition the error rate and confusion matrix are:
```{r, echo=FALSE }
model
```
The out-of-band error rate for the model is .76%.

For the testing partition the error rate and confusion matrix are:
```{r,echo=FALSE}
results <- predict(model,ptesting)
confusionMatrix(ptesting$classe,results)
```
This resulted an error rate on the testing partition of .51%.


Predict the outcomes for the testing data set:
```{r}
out <- as.character(predict(model,testing))
```
Testing set prediction = `r out`